---
title: "HW1 Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Example Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r }
library(bis557)
```
Use iris dataset for example
 
## Using linear_model()
Linear model function take in a dataset, a formula and an option contrast list to return to coefficients of the linear model
```{r}
form <- Sepal.Length ~ Sepal.Width
linear_model(formula = form,df = iris)
```

## Using gradient_descent()
Gradient descent function take in a dataset, a formula and an option contrast list to return to coefficients by using gradient descent, number of iterations and learning rate could also be modified

```{r}
form <- Sepal.Length ~ Sepal.Width
gradient_descent(formula = form,df = iris)
```

# HW1 
1.  Create an R package called `bis557`.
2.  Implement the `linear_model()` function and document it. It should
    take a formula, a data frame and a list of constasts for factor
    variables as input.
3.  Add a data.frame called `lm_patho` as a data set to the package. Use
    the `lm_path.csv` for the data. Create a `data-raw` directory in
    your package (hint: use `usethis::use_data_raw()` and
    `usethis::use_data()`) to add it to the package. Don’t forget to
    document it.
4.  Add the `test-linear-model.r` file to your regression testing (hint:
    use `usethis::use_testthat()` to create the directory structure).
    Make sure your implementation passes the tests.
5.  Implement gradient descent for ordinary least squares.
6.  Write test code for your gradient descent function.
7.  Add the package to [Travis-CI](https://travis-ci.com/) and update
    the `README.md` file so that the badge points to your Travis build
    (hint: use `usethis::use_travis()`). You will know it is working
    when the badge is green and gives the message “passing”. Extra hint:
    sign up for the Github student pack.
8.  Add code coverage (with either [Coveralls](https://coveralls.io/) or
    [CodeCov](https://codecov.io/)). Hint use `usethis::use_coverage()`).

